Cycles of the logistic map

### 2013-05-14 doc/pc-ijbc.tex
small doc changes

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx^{(#1)}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\Rlam}{(R, \lambda)}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$, $\ldots$
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = f(x_k) = x_{k+1}$;
the next horizontal arrow then reflects $(x_k, y)$ to
  $(y, y) = (x_{k+1}, x_{k+1})$,
  where the next iteration starts.
%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} or a constant (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle} or a self-repeating loop,
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

@@

Since the $n$-cycles constitute a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic} to \ref{sec:examples}).
The contributions from shorter $d$-cycles ($d|n$)
  are lower-degree factors of the polynomials,
  and can be readily removed (\refsec{primfac}).

At the first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \req{der}.
The result would contain $R$ only,
  and is therefore the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1}
  that gradually eliminates $x_k$ out of the $n+1$ equations.
As a special case, if $x_2, \dots, x_n$ are eliminated first,
  the final elimination of $x_1$ can be accomplished
  by setting the resultant of $f^n(x_1) = x_1$ and
  \refeq{der} or $\frac{d}{dx}f^n (x_1) = \pm 1$
  to zero \cite{burm},
  which furnishes the needed polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

Instead of solving \refeqs{logmaps} for $x_k$,
  we will derive a set of \emph{homogeneous linear} equations
  $\sum_j A_{ij} C_j = 0$
  regarding cyclic polynomials $\{C_j\}$ of $x_k$
  (an example of a cyclic polynomial is $x_1 x_2 + x_2 x_3 + \dots + x_n x_1$).
%
Now the matrix formed by the coefficients $A_{ij}$ of the equations
  must have a zero determinant,
  for the cyclic polynomials $\{C_j\}$ are not zeros altogether.
%
%As the coefficients are polynomials of $R$ only,
Thus, the zero-determinant condition $|\mathbf A| = 0$ gives
  the needed polynomial equation of $R$,
whose roots contain all fixed points of $f^n$.
%
%Note that we will obtain different equations
%  at the onset and bifurcation points
%  because the coefficients,
%  hence the matrices and determinants,
%  will be different.
%

@@

A cyclic polynomial is an invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ (for $n = 4$).
%
It is, however, not necessarily symmetric
  (i.e., invariant under the exchange of any two $x_k$ and $x_{j}$),
  e.g., $a$ and $b$ are not symmetric,
  but $a + b$ is.

@@

Cyclic polynomials
  generated from monomials of unit coefficient
  (such as $a$ and $b$, but not $a + b$ in the above example)
  can be labeled as follows.
We write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  and form a sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
The sequence $p$ serves as the label of the cyclic polynomial $C_p(\vxn{n})$,
  where $\vxn{n} \equiv \{x_1, \ldots, x_n\}$
[we will, however, drop the superscript ``$^{(n)}$'' or
  even the argument ``$(\vxn{n})$'' entirely when it is clear].
For example,
  $C_{12}(\vx)  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$ and
  $C_{112}(\vx) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%The cycle length $n$ is omitted for convenience.
%  for we will mostly work with a fixed $n$ at a time.
In case the cyclic polynomial has several generators,
%  e.g., both $x_1 x_2$ and $x_2 x_3$ are generators of $C_{12}(\vx)$
%   (assuming $n \ge 3$),
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ instead of $x_2 x_3$ for $C_{12}(\vx)$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0(\vx) \equiv 1$ for completeness.
%
%
%

@@

A cyclic polynomial is \emph{square-free}
  if it has no square or higher powers of $x_k$,
e.g., $C_{12}(\vx) = x_1 x_2 + \cdots$ is square-free,
  $C_{112}(\vx) = {x_1}^2 x_2 + \cdots$ is not.
%see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
The set of all square-free $p$ is denoted by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\dots n\}$,
and its size $|\Pset|$ equals the number $\NB$ of
  square-free cyclic polynomials.
%
%
The set of square-free cyclic polynomials
  furnishes a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx)$
  formed by the $n$-cycle points
  $\vx = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the square-free cyclic polynomials $C_p(\vx)$:
\[
  K(\vx) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx)$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since the original polynomial is cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx) \equiv 1$.
%Since no square or higher powers of $x_k$
%  can survive the reduction,
%  the final cyclic polynomials are square-free;
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vxn{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The matrix of the connecting coefficients then must have a zero determinant,
  which yields the desired polynomial.

@@

By \refeq{der},
  $\Lambda(\vx)$ equals to $\lambda = +1$ or $-1$
  at the onset or bifurcation point,
  respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx)$:
\begin{equation}
  \lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \lambda \, \vct I - \vct T(R) \big] \, \vct C = 0,
\tag{\ref{eq:xcp}'}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx)\}$ is an $\NB$-dimensional column vector.

Since a set of homogeneous linear equations
  has a non-trivial solution
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \big| \lambda \, \vct I - \vct T(R) \big| = 0.
\label{eq:xr}
\end{equation}
Here, $A_n\Rlam$ is a polynomial of $R$ and $\lambda$,
  and the subscript $n$ is attached such that
  it can be used later with $A_d\Rlam$ of divisors $d$ of $n$.
\refeq{xr} is a necessary condition
  since $C_p(\vx)$ cannot vanish altogether;
and since it involves $R$ only,
  the polynomial expansion of the determinant
  gives the answer to our problem.
%Note that we did not solve the values $C_p$ in the process.

To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$
  of the cycle points $\vx = \{x_k\}$
  are linear related by % \refeq{xcp},
  $\lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx)$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed point of $f^n$
  is marginally stable when $R$ satisfies
  characteristic equation $\big| \lambda \, \vct I - \vct T(R) \big| = 0$.
  \label{thm:main}
\end{theorem}

\begin{remark}
  Generally, the algorithm can trace a contour of marginal stability of $R$
  in the complex plane
  for complex cycles $\vx$
  if $\lambda$ is set to $\exp(i\theta)$ with $\theta \in [0, 2\pi]$.
  The region enclosed by the contour corresponds to
    a bulb in the inverted Mandelbrot set
  (it follows from the transformation $z_k = -x_k, c = -R$, which
  reduces \refeq{logmaps} to the quadratic map $z_{k+1} = c + {z_k}^2$
  used in the Mandelbrot set \cite{mandelbrot}),
  see ref. \cite{stephenson2, stephenson3}. % and \reffig{logbulb}.
\end{remark}

@@

We illustrate the above algorithm on several small-$n$ cases.
%
It is still helpful to have a mathematical software
  in order to verify certain steps
  (e.g., in computing the determinants and their factorization).

For $n = 1$, we have
  $C_0 = 1$, $C_1 = x_1$, $\Lambda = -2 \, x_1$, and
%Thus,
%$\Lambda \, C_0 = -2 \, x_1$,
%$\Lambda \, C_1 = -2 \, x_1^2 = -2 \, R + 2  \, x_1$,
%and
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
With $\Lambda \rightarrow \lambda$, the characteristic equation (\ref{eq:xr}) reads
\begin{equation}
0
  %=\left| \begin{array}{cc}
  %\lambda     & 2          \\
  %2 R         & \lambda - 2
  %\end{array} \right|
  = -4R - 2\lambda + \lambda^2.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
Thus, $R = (\lambda^2 -2\lambda)/4$;
  and the fixed point begins at $R_a = -1/4$ ($\lambda = +1$),
  and becomes unstable at $R_b = 3/4$ ($\lambda = -1$).

For $n = 2$, the cyclic polynomials
  are
$C_0 = 1$,
$C_1 = x_1 + x_2$,
$C_{12} = x_1 x_2$;
and $\Lambda = 4 \, x_1 x_2$.
Thus,
$\Lambda \, C_0 = 4 \, C_{12}$,
$\Lambda \, C_1 %= 4 \, {x_1}^2 \, x_2 + 4 \, x_1 \, {x_2}^2
  = -8 \, R \, C_0 + 4 \, (R + 1) \, C_1$
by the example after \refthm{sqrfree},
$\Lambda \, C_{12}
  = 4 \, (R - x_2) (R - x_1)$,
and
\[
\Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right)
 =
  \left( \begin{array}{ccc}
  0           & 0         & 4 \\
  -8R         & 4(R+1)    & 0 \\
  4R^2        & -4R       & 4
  \end{array} \right)
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right).
\]
%
%
%
The characteristic equation \refeq{xr} reads
%
%
%
\begin{align}
0 &= (4R - 4 + \lambda) [(4 R-\lambda)^2 - 4 \lambda].
  \tag{\ref{eq:xr}-2}
\label{eq:xr2}
\end{align}
%
%
%
The first factor is responsible for the 2-cycle,
  while the second factor is for unstable fixed points. % (see later).
Setting the former to zero yields $R = 1 - \lambda/4$;
so the onset ($\lambda = +1$) value is $R_a = 3/4$ ($r_a = 3$)
and the bifurcation ($\lambda = -1$) value is $R_b = 5/4$ ($r_b = 1+\sqrt 6$).
%
%Note that the onset point of the 2-cycle
%  is located at $R = 3/4$,
%  where the fixed point bifurcates \cite{strogatz}.
%  \big[compare \reffigs{cobweb}(a) and (b)\big].

@@

The characteristic polynomial $A_n\Rlam$ in \refeq{xr}
  is derived for the fixed points of $f^n$.
Thus, it represents not only the $n$-cycles,
  but also cycles of shorter periods $d$ ($d|n$).
%
We can remove the contributions from the shorter cycles by the following theorem.

>>>>

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx^{(#1)}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\Rlam}{(R, \lambda)}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$, $\ldots$
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = f(x_k) = x_{k+1}$;
the next horizontal arrow then reflects $(x_k, y)$ to
  $(y, y) = (x_{k+1}, x_{k+1})$,
  the starting point of the next iteration.
%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} or a constant (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle} or a self-repeating loop,
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

@@

Since the $n$-cycles points constitute a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The contributions from shorter $d$-cycles ($d|n$)
  are lower-degree factors of the polynomials,
  and can be readily removed (\refsec{primfac}).

At the first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \req{der}.
The result would contain $R$ only,
  and is therefore the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
As a special case, if $x_2, \dots, x_n$ are eliminated first,
  the final elimination of $x_1$ can be accomplished
  by setting the resultant of $f^n(x_1) = x_1$ and
  \refeq{der} or $\frac{d}{dx}f^n (x_1) = \pm 1$
  to zero \cite{burm},
  which furnishes the needed polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

Instead of solving \refeqs{logmaps} for $x_k$,
  we will derive a set of \emph{homogeneous linear} equations
  $\sum_j A_{ij} C_j = 0$
  regarding cyclic polynomials $\{C_j\}$ of $x_k$
  (an example of a cyclic polynomial is $x_1 x_2 + x_2 x_3 + \dots + x_n x_1$).
%
Now the matrix formed by the coefficients $A_{ij}$ of the equations
  must have a zero determinant,
  for the cyclic polynomials $\{C_j\}$ are not zeros altogether.
%
Thus, the zero-determinant condition $|\mathbf A| = 0$ gives
  the needed polynomial equation of $R$,
whose roots contain all fixed points of $f^n$.
%

@@

A cyclic polynomial is an invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ (for $n = 4$).
%
It is, however, unnecessarily symmetric
  (i.e., invariant under the exchange of any two $x_k$ and $x_{j}$),
  e.g., $a$ and $b$ are not symmetric,
  although $a + b$ is.

@@

Cyclic polynomials
  generated from monomials of unit coefficient
  (such as $a$ and $b$, but not $a + b$ in the above example)
  can be labeled as follows.
We write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  and form a sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
The sequence $p$ serves as the label of the cyclic polynomial $C_p(\vxn{n})$,
  where $\vxn{n} \equiv \{x_1, \ldots, x_n\}$
[we will, however, drop the superscript ``$^{(n)}$'' or
  even the argument ``$(\vxn{n})$'' entirely when it is clear].
For example,
  $C_{12}(\vx)  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$ and
  $C_{112}(\vx) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ instead of $x_2 x_3$ for $C_{12}(\vx)$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0(\vx) \equiv 1$ for completeness.
%
%
%

@@

A cyclic polynomial is \emph{square-free}
  if it has no square or higher powers of $x_k$,
e.g., $C_{12}(\vx) = x_1 x_2 + \cdots$ is square-free,
  $C_{112}(\vx) = {x_1}^2 x_2 + \cdots$ is not.
%see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
We denote the set of all square-free $p$ by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\dots n\}$,
and its size by $\NB$. 
% is the number of square-free cyclic polynomials.
%
%
The set of square-free cyclic polynomials
  furnishes a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx)$
  formed by the $n$-cycle points
  $\vx = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the square-free cyclic polynomials $C_p(\vx)$:
\[
  K(\vx) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx)$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement rules and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vxn{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The matrix of the connecting coefficients thus has a vanishing determinant,
  which furnishes the desired polynomial equation.

@@

By \refeq{der},
  $\Lambda(\vx)$ equals to $\lambda = +1$ or $-1$
  at the onset or bifurcation point,
  respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx)$:
\begin{equation}
  \lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \lambda \, \vct I - \vct T(R) \big] \, \vct C = 0,
\tag{\ref{eq:xcp}$'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx)\}$ is an $\NB$-dimensional column vector.

Since a set of homogeneous linear equations
  has a non-trivial solution
  (i.e., $C_p(\vx)$ do not vanish altogether)
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \big| \lambda \, \vct I - \vct T(R) \big| = 0.
\label{eq:xr}
\end{equation}
%
Now $A_n\Rlam$, after we expand the determinant,
  is a polynomial of $R$ and $\lambda$,
  the equation is the desired answer to our problem.
%
We attach the subscript $n$ to $A\Rlam$ because
  it will later be used with $A_d\Rlam$ of divisors $d$ of $n$.
%
To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$
  of the cycle points $\vx = \{x_k\}$
  are linearly related by % \refeq{xcp},
  $\lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx)$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed point of $f^n$
  is marginally stable when $R$ satisfies
  characteristic equation $\big| \lambda \, \vct I - \vct T(R) \big| = 0$.
  \label{thm:main}
\end{theorem}

%\begin{remark}
More generally, the algorithm can trace a contour of marginal stability of $R$
  in the complex plane
  for complex cycles $\vx$
  if $\lambda$ is set to $\exp(i\theta)$ with $\theta \in [0, 2\pi]$.
  The region enclosed by the contour corresponds to
    a bulb in the inverted Mandelbrot set
  (it follows from the transformation $z_k = -x_k, c = -R$, which
  reduces \refeq{logmaps} to the quadratic map $z_{k+1} = c + {z_k}^2$
  used in the Mandelbrot set \cite{mandelbrot}),
  see ref. \cite{stephenson2, stephenson3}. % and \reffig{logbulb}.
%\end{remark}

@@

We illustrate the above algorithm on several small-$n$ cases.
%
It is still helpful to have a mathematical software
  at hand to verify certain steps
  (e.g., in computing the determinants and their factorization).

For $n = 1$, we have
  $C_0 = 1$, $\Lambda = -2 \, C_1 = -2 \, x_1$, and
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
With $\Lambda \rightarrow \lambda$, the characteristic equation (\ref{eq:xr}) reads
\begin{equation}
0
  %=\left| \begin{array}{cc}
  %\lambda     & 2          \\
  %2 R         & \lambda - 2
  %\end{array} \right|
  = -4R - 2\lambda + \lambda^2.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
Thus, $R = (\lambda^2 -2\lambda)/4$;
  and the fixed point begins at $R_a = -1/4$ ($\lambda = +1$), and
  becomes unstable at $R_b = 3/4$ ($\lambda = -1$).

For $n = 2$, the cyclic polynomials
  are
$C_0 = 1$,
$C_1 = x_1 + x_2$,
$C_{12} = x_1 x_2$;
and $\Lambda = 4 \, x_1 x_2$.
Thus,
$\Lambda \, C_0 = 4 \, C_{12}$,
$\Lambda \, C_1 %= 4 \, {x_1}^2 \, x_2 + 4 \, x_1 \, {x_2}^2
  = -8 \, R \, C_0 + 4 \, (R + 1) \, C_1$
by the example after \refthm{sqrfree},
$\Lambda \, C_{12}
  = 4 \, (R - x_2) (R - x_1)$,
and
\[
\Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right)
 =
  \left( \begin{array}{ccc}
  0           & 0         & 4 \\
  -8R         & 4(R+1)    & 0 \\
  4R^2        & -4R       & 4
  \end{array} \right)
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right).
\]
%
%
%
The characteristic equation \refeq{xr} reads
%
%
%
\begin{align}
0 &= (4R - 4 + \lambda) [(4 R-\lambda)^2 - 4 \lambda].
  \tag{\ref{eq:xr}-2}
\label{eq:xr2}
\end{align}
%
%
%
The first factor is responsible for the 2-cycle,
  while the second factor is for the unstable fixed points. % (see later).
Setting the former to zero yields $R = 1 - \lambda/4$;
so the onset ($\lambda = +1$) value is $R_a = 3/4$ ($r_a = 3$)
and the bifurcation ($\lambda = -1$) value is $R_b = 5/4$ ($r_b = 1+\sqrt 6$).
%
%Note that the onset point of the 2-cycle
%  is located at $R = 3/4$,
%  where the fixed point bifurcates \cite{strogatz}.
%  \big[compare \reffigs{cobweb}(a) and (b)\big].

@@

The characteristic polynomial $A_n\Rlam$ in \refeq{xr}
  is derived for the fixed points of $f^n$.
Thus, it represents not only the $n$-cycles,
  but also cycles of shorter periods $d$ ($d|n$).
%
Fortunately, the contributions from different cycles
  are separate factors of $A_n\Rlam$.
For a large $n$,
  the factor $P_n\Rlam$ that represents all $n$-cycles
  can be easily recognized
  as the largest factor of $A_n\Rlam$.
%
More systematically, we have

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$,
  $\lambda^{1/c}$ is a complex $c$th root of $\lambda$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime.
%Thus,
The first few $\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$%.
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
\cite{hardy}.
%\end{remark}
%
%
\begin{remark}
%$B_{d,c}\Rlam$ is a polynomial of $\lambda$.
Despite the argument $\lambda^{1/c}$,
  the product $\prod_{k=1}^c A_d(R, e^{2k\pi i/c} \lambda^{1/c})$,
  hence $B_d^c\Rlam$,
  is free from radicals of powers of $\lambda^{1/c}$,
  for it is invariant under $\lambda \rightarrow e^{2\pi i} \lambda$.
Thus, $B_d^c\Rlam$ is a polynomial of $\lambda$,
and $\deg_\lambda B_d^c\Rlam = \deg_\lambda A_d\Rlam$.
Particularly, $B_n^1\Rlam = A_n\Rlam$.
\end{remark}
%
%
%We will refer to $P_n\Rlam$ as the \emph{primitive factor} below.
%
%
Let us apply \refeq{primfac}
to the examples in \refsec{examples}.
We will omit the argument $R$ below for convenience.
%%
For $n = 1$,
%there is no irrelevant factor in \refeq{xr1} and
$P_1(\lambda) = B_1^1(\lambda) = A_1(\lambda)
  = \lambda^2 - 2\,\lambda - 4\,R$.
%
%
%
For $n = 2$, since
$B_1^2(\lambda)
=
A_1(\sqrt{\lambda})
A_1(-\sqrt{\lambda})
=
(4R-\lambda)^2 -4\lambda$,
%
$P_2(\lambda) = A_2(\lambda) \, B_1^2(\lambda)^{-1} = 4R - 4 + \lambda$.
%
%
For $n = 4$,
%we have
%$16 (R-1)^2 - \lambda
%= (4R - 4 + \sqrt{\lambda})(4R - 4 - \sqrt{\lambda})$
%and
%$
%(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda
%=
%\big[(4R - \sqrt{\lambda})^2  - 4\sqrt{\lambda}\,\big]
%\big[(4R + \sqrt{\lambda})^2  + 4\sqrt{\lambda}\,\big]$.
%Thus,
we can show that
the last two factors of \refeq{xr4}
can be written as
$B_2^2(\lambda) = A_2(\sqrt{\lambda}) A_2(-\sqrt{\lambda})$,
and
%
$P_4(\lambda)
    = A_4(\lambda) \, B_2^2(\lambda)^{-1}$
indeed the yields the sextic factor.
%$    = 4096 R^6-12288 R^5+ 256 (\lambda + 48) (R^4- R^3)
%    -16(\lambda + 32)(\lambda - 16) R^2 - (\lambda-16)^3$.
Note, $B_1^4(\lambda)$ is unused for $\mu(4) = 0$.

%It can be, however, problematic,
%   if $P_n(R, \lambda)$ is solved for $\lambda = 1$
%   instead of a general $\lambda$,
%   for $P_n(R, 1)$ itself can be further factorized.
%
Due to the technical nature of the derivation and subsequent discussions,
  the reader may wish to skip the rest of \refsec{logmap}
  on first reading.
